Using SMT engine to generate Symbolic Automata\thanks{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai}

### 2018-01-16 2017-Xudong-SMT/FM-2018-extended.tex
New version of the BIP diagramme for the Failure Monitor architecture -- consumes less space on the page Yet more cleanup in the .gitignore file
up to page 11

%% This paper presents our open automaton computing algorithm and its
%% implementation using the pNet API from the VerCors platform.
%% Comparing
%% the result of the algorithm with 
%% the manual result on our running examples shows that the algorithm
%% works properly in computing the pNet semantics.
%% The
%% algorithm contains the method computing the open transitions composing
%% the transitions from the subnets according to several defined rules
%% the open transitions need to satisfy. These results are not the final
%% results in fact, because they contain some unsatisfiable results we need
%% to eliminate. To eliminate those unsatisfiability, we interact with an
%% automatic theorem prover for satisfiability checking.

\paragraph{Contribution}
In the article we show how:
\begin{itemize}
\item We define the open automaton generation algorithm, and we
  implemented a full working prototype, within the 
  VerCors platform. In the process, we improved the
  semantics rules from~\cite{henrio:Forte2016}, and add features in
  the algorithm to deal the full 
  model, including management of variables and assignments.
%  \item We extended the meta-model of action expressions, and the
%    associated API, in the VerCors platform, to be able to manage
%    parameterisation over the action algebra operators.
\item We implement the interaction between our algorithm and the Z3
    SMT solver, for checking satisfiability of the transitions
    generated by the algorithm.
  \item \ERIC{We show the interest of this approach on an
    industry-inspired case-study, namely one architectural pattern
    extracted from the BIP specification of a nanosatellite on-board
    software.} 
\end{itemize}

@@

For other applications, e.g. for the analyses of 
programming languages, there exist dedicated platforms making use of
external automatic theorem 
provers (ATP), or even some automatic tactics from interactive theorem
provers (ITP), to perform symbolic reasoning, and for example to
discharge some subgoals in the proofs.
Tools like Rodin~\cite{deharbe2013,deharbe2014,abrial2007} have
already integrated several provers, like Z3, as modules for proving
the proof obligations generated from the model provided by the user. 
The prover we used, which also happens to be Z3, is developed by Microsoft Research
based on the satisfiability modulo 
theories (SMT), mainly applied in extended static checking, test case
generation, and predicate abstraction.
In a similar way, there are several ATPs/ITPs we could consider to use for
the result pruning and bisimulation checking in our algorithm, as an
alternative to Z3, such as CVC4~\cite{barrett:CAV2011},
Coq~\cite{armand:CPP2011},
%Coq~\cite{armand:CPP2011,besson:CPP2011},
Isabelle~\cite{blanchette:FroCoS2011} or others.

\paragraph{Structure.}
In section
\ref{section:pnets} we give a description and a formal definition of
the pNet model, as found in previous publications. Then in
\ref{section:examples} we present a simple example that will
illustrate the rest of the paper.
Section \ref{section:op-semantics} recalls the operational semantics
of pNet, including its structural rules.
%, and a sketch of the algorithm
%to compute such semantics, but also explains the enhancement we have
%done to these definitions
Section \ref{section:implementation} explains in details our
implementation within the Vercors platform, and shows the full result of
the semantic computation on the running example.
Finally we conclude and discuss perspectives in section
\ref{section:conclusion}.

@@

%\smallskip\noindent
\paragraph*{Notations.}
We extensively use indexed structures
over some countable indexed sets, which are equivalent to mappings over
the countable set. % . The indexes will usually be
% integers, bounded or not. Such an indexed family is
%denoted
%follows:
$a_i^{i\in I}$
%, or equivalently  $(i\mapsto a_i)^{i\in I}$
denotes a family of elements $a_i$ indexed over the
set $I$. % Such a family
% is equivalent to the mapping $(i\mapsto a_i)^{i\in I}$.
% To specify the set over which the structure is indexed,
% indexed structures are always denoted with an exponent of the form $i\in I$
% (arithmetic only appears in the indexes if necessary).
$a_i^{i\in I}$ defines both $I$ the set over which the family is
indexed (called \emph{range}), and $a_i$ the elements of the family.
E.g., $a^{i\in\{3\}}$ is the mapping with a single entry $a$ at index
$3$ ; abbreviated $(3\mapsto a)$ in the following.
When this is not
ambiguous, we shall use notations for sets, and typically write
``indexed set over I'' when formally we should speak of multisets, and
write $x\in a_i^{i\in I}$ to mean $\exists i\in I.\, x=a_i$.  An empty
family is denoted $\emptyset$. We
denote classically $\overline{a}$ a family when the indexing set is
irrelevant.  $\uplus$ is the disjoint union on
indexed sets.

\TODO{Eric: merge term algebra, algebra pres, and add a short example (BIP)}

\paragraph*{Term algebra.}
Our models rely on a notion of parameterized actions, that are
symbolic expressions using data types and variables. As our model aims
at encoding the low-level behavior of possibly very different
programming languages, we do not want to impose one specific algebra
for denoting actions, nor any specific communication mechanism. So we
leave unspecified the constructors of the algebra that will allow building
expressions and actions. Moreover, we use a generic {\em action interaction}
mechanism, based on unification between two or more action
expressions. This will be used in the semantics of synchronization
vectors to express various kinds of communication or synchronization mechanisms.

@@

pNets can encode naturally the notion of input actions as found e.g. in value-passing CCS
\cite{Milner89} or of usual point-to-point message passing calculi, but it also allows
for more general mechanisms, like gate negotiation in Lotos, or broadcast
communications.
\QIN{
\paragraph*{Algebra presentations}
\begin{definition}
  An Algebra Presentation is a pair $\mathcal{P}=<Sorts,Constrs,Ops>$ where:
  \begin{itemize}
  \item $Sorts$ is a set of Sort names
    \footnote{Later we may want to extend this with Sort constructors,
      like Array or Pair, but this is not needed now}
  \item $Constrs$ is a set of constructor operators, with $Con$ the
    constructor name, and arity: $arity(Con)=n \in \mathbb{N}$,
    and with their signature and their associated selectors,
    of the form: '$Con : (sel_1,sort_1), ... , (sel_n,sort_n) -> sort$'.
    For each argument, the pair $(sel_i,sort_i)$ defines an auxiliary
    operator of name $sel_i$ with signature $sel_i : sort -> sort_i$.
    \item $Ops$ is a set of other (auxiliary) operators, with their
      arity and signature, of the form: $Op : sort_1, ...  sort_n ->
      sort$
      \item $Constrs(sortname), Sels(sortname)$, respectively define the set of
        constructors and selectors of sort $sortname$
  \end{itemize}
  All sorts and operator names must be distinct.\\
  Amongst Constructors, those of arity 0 are called constants, and we
  define $Consts(\mathcal{P}) = \{Con \in Constrs. arity(Con)=0\}$.
\end{definition}
}

@@

\begin{figure}[t]
   \centerline{\includegraphics[width=12cm]{XFIG/FailureTimerPNET}}
   \caption{FailureTimer pNet encoding}  \label{schema:BIParchitecture}
 \end{figure}

@@

The semantics of open pNets will be defined as an open automaton,that
is an automaton where each transition composes transitions 
of several LTSs with the
actions of some holes; the transition occurs if some predicates hold,
and can involve a set of state modifications.

@@

\paragraph{Structural Semantic Rules:}
\TODO{I have pushed the rules in the appendix, we need to add a short high
  level desciption here.}

\TODO{ERIC: I choose here to remove the expample of construction of an
  OT as a deduction tree. First because it was already published in
  the Forte'16 paper, but also because it is not the way the algorithm
  is working.}

%% \begin{example} \emph{Using the operational rules to compute  open-transitions: }
%%   in Fig. \ref{usingrules:ot2} we show the deduction tree used to construct and prove the 
%%   open transition $ot_2$.
%% \end{example}

%% \begin{figure}[h]
%% %  \vspace{-10mm}

%% \begin{mathpar}
%%   \small
%%   \inferrule
%%     {
%%           \inferrule
%%               {0 \xrightarrow {\delta}_{C_1} 0 ~~
%%                [ s_0=0 ]
%%               }
%%               {C_1
%%                 \models
%%                 \openrule
%%                       {0 \xrightarrow{\delta}_{C_1} 0 ~~
%%                        [ s_0=0 ] ~~
%%                        \{ s_0 \leftarrow 1\}
%%                       }
%%                       {\ostate{0}\xrightarrow{\delta}\ostate{0}}
%%               }
%% 	\\
%%       \inferrule
%%         {
%%           \inferrule
%%               {0 \xrightarrow{acc(x)}_{C_2} 1}
%%               {C_2
%%                 \models
%%                 \openrule
%%                       {0 \xrightarrow{acc(x)}_{C_2} 1}
%%                       {\ostate{0}\xrightarrow{acc(x)}\ostate{1}}
%%               }
%%         }
%%         {
%%           \textrm{acc(x);Q}\models
%%               \openrule
%%                   { 0 \xrightarrow {acc(x)}_{C_2} 1}
%%                   {\ostate{0}\xrightarrow{acc(x)}\ostate{1}}
%%         }   
%%     }
%%     {
%%      \textrm{PN1}
%%      \models
%%         \openrule{
%%                         0 \xrightarrow{\delta}_{C_1} 0, 0 \xrightarrow{acc(x)}_{C_2} 1 ~~
%%                        \{\xrightarrow{hb_{12}}_P\} ~~ 
%%                         [s_0=0 \!\land\! hb_{12}=\delta(x)] ~~ 
%%                        \{s_0 \xleftarrow{} 1\}           
%%                       }
%%      {\ostate{00} \xrightarrow{\tau} \ostate{01}}
%%       }\vspace{-4ex}
%% \end{mathpar}
%%   \caption{Proof of  $ot_{2}$ (interaction of process $P$ and action $acc(x)$) for ``P>>(acc(x);Q)''}
%%   \label{usingrules:ot2}
%% \end{figure}

%% The proof tree uses TR1 twice, for the $\delta$ transition of $C_1$
%% and for the $acc(x)$ transition of $C_2$, then uses an action
%% $hb_{12}$ of hole $P$, and combines the results using the first vector
%% of the PN2 sub-pNet, and the second vector of the top node according
%% to TR2. This yields a final $\underline{\delta(x)}$ transition. The
%% deduction tree in the figure shows how the predicates are generated in
%% this process.

@@

@@

\begin{algorithm}
\caption{Open Automaton Generation}
\begin{algorithmic}[1]
\Require The pNet node P.
\State Initialize sets $U$, $E$ for unexplored/explored global states, $L$ for result OTs;
\While{!isEmpty(U)}
	\State Chose $S$ in $U$;
	\State OT = MakeTransitions(P, S);
        \State Store $OT$ in $L$; 
	\For{each target global state $T$ from $OT \in L$}
		\IfThen{$(!U.contains(T))\ \&\&\ (!E.contains(T))$}{Add $T$ into the $U$;}
	\EndFor
\EndWhile
\State Prune the open transitions in $L$ using the SMT solver;
\State \Return $L$;

@@

In a pLTS, there may be several incoming transitions of some states
that assign potentially different values to a state variable.
To handle such cases, the algorithm manages the variables together
with a list for each pLTS.
The list we used contains several triples <$v$, $S$, $AssignRH$> 
where $v$ is the variable in pLTS, $S$ is its owner state and
$AssignRH$ is a list of expressions over other variables in the right
hand side of assignments of $v$.
As a pragmatic extension to the formal definition, we also manage
``global variables'', defined in all states of the pLTS.

@@

We use a brute-force method to generate all the possible open
transitions in the open automaton, using all possible combinations of
synchronization vectors. Naturally this builds some transitions where
the predicates express incompatible constraints. Even if having an
unsatisfiable (symbolic) transition in the open automaton would not be incorrect,
we want to check them for satisfiability, and reduce the number of
transitions and states in the automaton.
\TODO{Change...} In Fig.~\ref{schema:unsat-ot}, we display an example of an
unsatisfiable open transition 
from the result of our example. It shows the case where
the controller C1 wants to move the control from $P$ to $(acc(x);Q)$,
conducting a $\delta$ transition. However, a synchronization
vector is chosen that does not match with this action, and we can
easily find the contradiction in the 
generated predicate term ``$\OTvar{:ra:11:1}=l \!\OTland\! \delta=\OTvar{:ra:11:1}$''.

@@

As seen in section
\ref{section:pnets}, the \emph{term algebra} used to express the
action expressions and data parameters is left open in
the pNet definition.

@@

\section{pNet semantic rules}

\paragraph{Note:} For convenience, we have included here the semantic rules for pNets,
that were already published in \cite{Forte}.

@@

\section{More details on the generation algorithm}
\subsection{Algebra specification}
\subsection{Term algebra: type system and static semantics}
Our models rely on a notion of parameterized actions, that are
symbolic expressions using data types and variables. As our model aims
at encoding the low-level behavior of possibly very different
programming languages, we do not want to impose one specific algebra
for denoting actions, nor any specific communication mechanism. So we
leave unspecified the constructors of the algebra that will allow building
expressions and actions. Moreover, we use a generic {\em action interaction}
mechanism, based on unification between two or more action
expressions. This will be used in the semantics of synchronization
vectors to express various kinds of communication or synchronization mechanisms.

\subsection{Fresh variables}

For example, in the running example, the first possible behavior of
the hole, it doesn't have a name, only have the prefix ``hb''. The
hole is the second subnets of the root node so its ``tree index'' is
``12''. The fresh variable name is ``$\OTvar{:hb:12:1}$''. 
\subsection{Variable managment}
The variables in each synchronisation vector are considered local: for a given pNet expression, we must have fresh local variables for each occurrence of a vector (= each time we instantiate rule Tr2). Similarly the state variables of each copy of a given pLTS in the system, must be distinct, and those created for each application of Tr2 have to be fresh and all distinct. This will be implemented within the open-automaton generation algorithm, e.g. using name generation using a global counter as a suffix.

@@

Note that in principle, such an algebra correspond to a given
high-level language (e.g. a process algebra), and that the algebra
presentation will be defined once and for all in the framework of the
pNet semantics of each specific language.

\paragraph{Algebra presentations}
\begin{definition}
  An Algebra Presentation is a pair $\mathcal{P}=<Sorts,Constrs,Ops>$ where:
  \begin{itemize}
  \item $Sorts$ is a set of Sort names
    \footnote{Later we may want to extend this with Sort constructors,
      like Array or Pair, but this is not needed now}
  \item $Constrs$ is a set of constructor operators, with $Con$ the
    constructor name, and arity: $arity(Con)=n \in \mathbb{N}$,
    and with their signature and their associated selectors,
    of the form: '$Con : (sel_1,sort_1), ... , (sel_n,sort_n) -> sort$'.
    For each argument, the pair $(sel_i,sort_i)$ defines an auxiliary
    operator of name $sel_i$ with signature $sel_i : sort -> sort_i$.
    \item $Ops$ is a set of other (auxiliary) operators, with their
      arity and signature, of the form: $Op : sort_1, ...  sort_n ->
      sort$
      \item $Constrs(sortname), Sels(sortname)$, respectively define the set of
        constructors and selectors of sort $sortname$
  \end{itemize}
  All sorts and operator names must be distinct.\\
  Amongst Constructors, those of arity 0 are called constants, and we
  define $Consts(\mathcal{P}) = \{Con \in Constrs. arity(Con)=0\}$.
\end{definition}

>>>>

\paragraph{Contribution}
In the article we show how:
\begin{itemize}
\item We define the open automaton generation algorithm, and we
  implemented a full working prototype, within the 
  VerCors platform. In the process, we improved the
  semantics rules from~\cite{henrio:Forte2016}, and add features in
  the algorithm to deal the full 
  model, including management of variables and assignments.
\item We implement the interaction between our algorithm and the Z3
    SMT solver, for checking satisfiability of the transitions
    generated by the algorithm.
\item We show the interest of this approach on an
    industry-inspired case-study, namely one architectural pattern
    extracted from the BIP specification of a nanosatellite on-board
    software.
\end{itemize}

@@

For other applications, e.g. for the analyses of 
programming languages, there exist dedicated platforms making use of
external automatic theorem 
provers (ATP), or even some automatic tactics from interactive theorem
provers (ITP), to perform symbolic reasoning, and for example to
discharge some subgoals in the proofs.
Tools like Rodin~\cite{deharbe2013,deharbe2014,abrial2007} have
already integrated several provers, like Z3, as modules for proving
the proof obligations generated from the model provided by the user. 
The prover we used, which also happens to be Z3, is developed by Microsoft Research
based on the satisfiability modulo 
theories (SMT), mainly applied in extended static checking, test case
generation, and predicate abstraction.
In a similar way, there are several ATPs/ITPs we could consider to use for
the result pruning and bisimulation checking in our algorithm, as an
alternative to Z3, such as CVC4~\cite{barrett:CAV2011},
Coq~\cite{armand:CPP2011},
Isabelle~\cite{blanchette:FroCoS2011} or others.

\paragraph{Structure.}
In section
\ref{section:pnets} we give a description and a formal definition of
the pNet model, as found in previous publications. Then in
\ref{section:examples} we present a simple example that will
illustrate the rest of the paper.
Section \ref{section:op-semantics} recalls the operational semantics
of pNet, including its structural rules.
Section \ref{section:implementation} explains in details our
implementation within the Vercors platform, and shows the full result of
the semantic computation on the running example.
Finally we conclude and discuss perspectives in section
\ref{section:conclusion}.

@@

%\smallskip\noindent
\paragraph*{Notations.}
We extensively use indexed structures
over some countable indexed sets, which are equivalent to mappings over
the countable set. % . The indexes will usually be
% integers, bounded or not. Such an indexed family is
%denoted
%follows:
$a_i^{i\in I}$
denotes a family of elements $a_i$ indexed over the
set $I$.
When this is not
ambiguous, we shall use notations for sets, and typically write
``indexed set over I'' when formally we should speak of multisets, and
write $x\in a_i^{i\in I}$ to mean $\exists i\in I.\, x=a_i$.  An empty
family is denoted $\emptyset$. We
denote classically $\overline{a}$ a family when the indexing set is
irrelevant.  $\uplus$ is the disjoint union on indexed sets.

\paragraph*{Term algebra.}
Our models rely on a notion of parameterized actions, that are
symbolic expressions using data types and variables. As we want to encode
the low-level behavior of possibly very different
programming languages, we do not want to impose one specific algebra
for denoting actions, nor any specific communication mechanism. So we
leave unspecified the constructors of the algebra that will allow building
expressions and actions. Moreover, we use a generic {\em action interaction}
mechanism, based on unification between two or more action
expressions. This will be used in the semantics of synchronization
vectors to express various kinds of communication or synchronization mechanisms.

@@

\paragraph*{Algebra presentations.}
In practice, the parameterization of the pNet model by some specific
action is realized through the definition of an ``algebra
presentation''. It will be used to define formaly a notion of
well-formness of a pNet system, and the translation of the pNet
semantics into the SMT engine input language.

\begin{definition}
  An Algebra Presentation is a pair $\mathcal{P}=<Sorts,Constrs,Ops>$ where:
  \begin{itemize}
  \item $Sorts$ is a set of Sort names
    \footnote{Later we may want to extend this with Sort constructors,
      like Array or Pair, but this is not needed now}
  \item $Constrs$ is a set of constructor operators, with $Con$ the
    constructor name, and arity: $arity(Con)=n \in \mathbb{N}$,
    and with their signature and their associated selectors,
    of the form: '$Con : (sel_1,sort_1), ... , (sel_n,sort_n) -> sort$'.
    For each argument, the pair $(sel_i,sort_i)$ defines an auxiliary
    operator of name $sel_i$ with signature $sel_i : sort -> sort_i$.
    \item $Ops$ is a set of other (auxiliary) operators, with their
      arity and signature, of the form: $Op : sort_1, ...  sort_n ->
      sort$
      \item $Constrs(sortname), Sels(sortname)$, respectively define the set of
        constructors and selectors of sort $sortname$
  \end{itemize}
  All sorts and operator names must be distinct.\\
  Amongst Constructors, those of arity 0 are called constants, and we
  define $Consts(\mathcal{P}) = \{Con \in Constrs. arity(Con)=0\}$.
\end{definition}

@@

\begin{figure}[t]
   \centerline{\includegraphics[width=12cm]{XFIG/FailureTimerPNET}}
   \caption{FailureTimer pNet encoding\TODO{on peut gagner un peu en hauteur}}  \label{schema:BIParchitecture}
 \end{figure}

@@

The semantics of open pNets will be defined as an open automaton, that
is an automaton where each transition composes transitions 
of several LTSs with the
actions of some holes; the transition occurs if some predicates hold,
and can involve a set of state modifications. Each state of an open
automaton has a set of \emph{state variables} that can be assigned
incoming transitions.

@@

\paragraph{Structural Semantic Rules:}
The semantics of pNet in term of open automata has been defined in
\cite{henrio:Forte2016}. For convenience we include these 
rules in Appendix \ref{appendix:semRules}.

@@

\begin{algorithm}[h]
\caption{Open Automaton Generation}
\begin{algorithmic}[1]
\Require The pNet node P.
\State Initialize sets $U$, $E$ for unexplored/explored global states, $L$ for result OTs;
\While{!isEmpty(U)}
	\State Chose $S$ in $U$;
	\State OT = MakeTransitions(P, S);
        \State Store $OT$ in $L$; 
	\For{each target global state $T$ from $OT \in L$}
		\IfThen{$(!U.contains(T))\ \&\&\ (!E.contains(T))$}{Add $T$ into the $U$;}
	\EndFor
\EndWhile
\State Prune the open transitions in $L$ using the SMT solver;
\State \Return $L$;

@@

@@

In a pLTS, there may be several incoming transitions 
that assign potentially different values to a state variable.
To handle such cases, the algorithm manages the assignments
within a list for each pLTS.
The list we used contains triples <$v$, $S$, $AssignRH$> 
where $v$ is the variable in pLTS, $S$ is its owner state and
$AssignRH$ is a list of expressions over other variables in the right
hand side of assignments of $v$.
% As a pragmatic extension to the formal definition, we also manage
% ``global variables'', defined in all states of the pLTS.
For a global state in the open automaton, getting the set of state
variables (which may be used in a transition) will simply be the union
of state variables of the individual pLTS states constituting this
global state (variables of different pLTS are distinct). We define:
\begin{itemize}
  \item $SVars(gs)$ the state variables of all states in the global
    state $gs$
  \item $Assigns(svar)$ the set of possible assignments of variable
      $svar$.
\end{itemize}

@@

Our matching/filtering strategy builds some transitions where
the predicates express incompatible constraints. Even if having an
unsatisfiable (symbolic) transition in the open automaton would not be incorrect,
we want to check them for satisfiability, and reduce the number of
transitions and states in the automaton.
In Fig.~\ref{schema:unsat-ot}, we display an example of an
unsatisfiable open transition 
from the result of our example. It shows the case where
the failure controller performs a ``$resume$'' action, while the timer
executes a ``$start$''. The chosen synchronization
vector does not match with these actions,as appears in the predicate
fragment ``$\OTvar{resume(false)}=\OTvar{fail(}\OTvar{b}_1\OTvar{)}$''.
Fig.~\ref{schema:smt-lib} shows the translation of this OT predicate
in the SMTlib syntax. Here we display also the proof generated by Z3,
though in practice our tool only needs the brute information ``sat'' or ``unsat''.

@@

%% As seen in section
%% \ref{section:pnets}, the \emph{term algebra} used to express the
%% action expressions and data parameters is left open in
%% the pNet definition.

%% So, for a given language, we now need to specify its data and action
%% domains, giving them an abstract syntax (sorts, constructors, operators,
%% and predicates). We call this a \emph{Presentation}.
%% We also define a concrete syntax, that will be used for
%% pretty-printing, but also for translating the presentation, and the
%% predicates, into Z3 syntax (see the ``declare-datatypes'' statement in
%% Fig~\ref{schema:smt-lib}).

%% An algebra presentation contains:

%% \begin{itemize}
%% 	\item Sorts: Constants sets of the algebra or types of the data
%%           and actions (integer and boolean sort are implicitly
%%           declared).  There \emph{must} be one sort for actions, and
%%           eventually others for the types of data
%%           parameters. In the LOTOS example, 
%%           we have the $Action$ sort: $\OTvar{Action = } \{ l, \delta,
%%           r, p, q, acc \} .$

%% 	\item Operators:
%%           %Operators used to construct action expressions; they are
%%           constructors and predicates for data and action sorts. For
%%           instance, in our LOTOS example, we
%%           have a generic action constructor ``$ACT$'', taking as
%%           argument an action litteral and a data parameter, declared as:
%%           $\OTvar{ACT:}\OTvar{< Action, Data >} \rightarrow \OTvar{Action.}$, 
%%         \item Constants: The satisfiability solver need to know which
%%           actions in the pLTS behaviors are constants (like the $l$, $\delta$, 
%%           $r$ of the controllers in running example). 
%%           For a given pNet system, the sort $Action$ will include all
%%           these constants.
%% \end{itemize}

@@

\newpage
\appendix
\section{Appendix}
The following appendices are included for the convenience of reviewers.
Our intention is to publish as a research report, before final publication, an
extended version containing all technical detials, proofs, and more examples.

\section{pNet semantic rules}
\label{appendix:semRules}

\paragraph{Note:} For convenience, we have included here the semantic rules for pNets,
that were already published in \cite{henrio:Forte2016}.

@@

\subsection{Variable managment}
The variables in each synchronisation vector are considered local: for a given pNet expression, we must have fresh local variables for each occurrence of a vector (= each time we instantiate rule Tr2). Similarly the state variables of each copy of a given pLTS in the system, must be distinct, and those created for each application of Tr2 have to be fresh and all distinct. This will be implemented within the open-automaton generation algorithm, e.g. using name generation using a global counter as a suffix.

@@

Note that in principle, an action algebra corresponds to a given
high-level language (e.g. a process algebra), and that the algebra
presentation will be defined once and for all in the framework of the
pNet semantics of each specific language.
